---
title: "Potential asymmetries in Arabis F2 reproductive success"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Background*: Crosses were performed between Fr2 plants (France2), which are self-compatible and have very low scent emission, and plants from multiple Italian populations, which are self-incompatible and have variable scent profiles across populations but generally high emission rates.  These F1s were then selfed to create F2s.


**Table 1**. A summary of the parentage of the F1s that were selfed to make the F2s. Maternal and paternal seed sources refer to the parental generation.

Maternal seed source | Paternal seed source(s)
----------------------|----------------------
Fr2 | It2, It5, It6, It8, It9, It10, It17
It2 | Fr2
It6 | Fr2
It8 | Fr2
It9 | Fr2 

The crosses that we have represented in the F2s in both directions are Fr2 with It2, It6, It8, and It9.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(knitr)
library(lme4)
library(lmerTest)
library(emmeans)

seeds <- read_csv("F1_pollinations_2021.csv")

ids <- read_csv("F1_individuals_2020.csv")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
#Add ID information to the seed counts:
seed_ids <- left_join(seeds, ids, by=c("Mother_ID"="Full_ID"))
```

**Table 2**: A summary of seed production by F1 cross type. `Cross_Type` is mother population x father population. `N_Cross` is the number of crosses of unique crosses of the cross type that were performed (e.g., different parental plants from the populations). `N_F1` is the total number of F1 plants from a given cross that were selfed to make the F2s. `N_Fruits` is the total number of fruits from that cross (e.g., the sum of all of the fruits from each of the plants from the particular cross type). `Seed_Count` is the total number of seeds generated from all of the fruits and plants of the particular cross type. `SeedsperFruit` is the average number of seeds in a fruit, to help identify which plants produced fruits with more seeds and compare across crosses that had different number of plants or fruits. 

```{r, warning=FALSE, message=FALSE, echo=FALSE}
counts <- seed_ids %>% filter(Treatment=="Self") %>% mutate(N_Seeds = replace_na(N_Seeds, 0)) %>% group_by(Cross_Type) %>%  summarize(N_Cross=length(unique(ID)), N_F1=length(unique(Mother_ID)), N_Fruits=length(Mother_ID), Seed_Count=sum(N_Seeds), SeedsperFruit=round(mean(N_Seeds), digits=0)) %>% filter(Seed_Count>0)

kable(counts)
```

*Observation*: Aside from the IT6 x Fr2 cross, which set a lot of seeds in both directions, in general it seems like crosses with Italian mothers set fewer seeds overall and fewer seeds per fruit.


To test if reproductive success is asymmetric depending on the cross direction, I analyzed whether reproductive success measures (seeds per fruit, total seed set) differ between crosses with a French mother vs crosses with an Italian mother. The specific parents used to create a particular F2 (because sometimes there were multiple F1s from a specific cross, see `N_Cross` in Table 2 above) was nested within the Italian population involved in the cross (because the French population was always Fr2) and this was included as a random effect in the model.

```{r, warning=FALSE, message=FALSE, echo=FALSE}

obj <- seed_ids %>% filter(Treatment=="Self") %>% mutate(N_Seeds = replace_na(N_Seeds, 0), F1_ID=Mother_ID) %>% group_by(Cross_Type, F1_ID) %>% summarize(N_Fruits=length(Mother_ID),Seed_Count=sum(N_Seeds), SeedsperFruit=mean(N_Seeds)) %>% filter(!is.na(Cross_Type)) %>% mutate(Cross_Type2 =if_else(str_detect(Cross_Type,"^Fr"), "Fr", "It")) %>% separate(F1_ID, c("Cross", "Individual")) %>% separate(Cross_Type, c("Mother", "Father"), sep="x") %>% mutate(Italian_Pop = if_else(str_detect(Mother, "^IT"), Mother, "NA")) %>% mutate(Italian_Pop = if_else(str_detect(Italian_Pop, "NA"), Father, Italian_Pop))


model <- lmer(SeedsperFruit~Cross_Type2+(1|Italian_Pop/Cross), data=obj)

#par(mfrow=c(1,2))
#hist(resid(model))
#plot(predict(model),resid(model)) ; abline(h=0)

```

Here is the anova output for the model of seeds/fruit:
```{r, warning=FALSE, message=FALSE, echo=FALSE}
anova(model)
```

And the numerical results of the contrast between the groups and a plot of the estimated marginal means:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
pairs(emmeans(model, "Cross_Type2"))
plot(emmeans(model, "Cross_Type2")) + xlab("Seeds/fruit (estimated marginal mean)") + theme_classic()+scale_y_discrete("Country of the mother", c("Fr", "It"), c("France", "Italy"))
```

*Crosses with French mothers set more seeds than crosses with Italian mothers.* Analyzing total seed set rather than seeds per fruit gives the same result.


